---
title: "Analysis of Seed and Fruit Data"
author: "Ana Belem Garcia, Arfa Burghias, Cann Doan, Fatima Tahir"
date: "2024-06-16"
format: html
editor: visual
---

## Introduction

This report analyzes the relationship between seed count, fruit diameter, and treatment effects on fruit development. We generate visualizations, statistical tests, and interpret results to understand seed distribution, fruit characteristics, and treatment effects.

## Load Required Libraries and Data

```{r setup, include=FALSE, echo=TRUE, fig.width=8, fig.height=6, fig.show='hold'}
# Load necessary librarie
library(ggplot2)
library(dplyr)

# Load the dataset
data <- read.csv("DATA/frutasysemillas.csv")
knitr::opts_chunk$set(fig.path = 'results/')
```

## Explanation

```         
•   Load ggplot2 for data visualization and dplyr for data manipulation.

•   Load the dataset named frutasysemillas.
```

## 1. Frequency Distribution of Number of Seeds

```{r}
#| label: Frequency distribution
ggplot(data, aes(x = Nseeds)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Frequency Distribution of Number of Seeds",
       x = "Number of Seeds",
       y = "Frequency") +
  theme_minimal()
```

## Explanation

```         
•   A histogram shows the frequency distribution of number of seeds per fruit.

•   binwidth = 5 groups the seeds into intervals of 5.
```

## 2. Density Curve of Number of Seeds

```{r}
#| label: Density Curve of NSeeds
ggplot(data, aes(x = Nseeds)) +
  geom_density(fill = "lightblue", alpha = 0.7) +
  labs(title = "Density Curve of Number of Seeds",
       x = "Number of Seeds",
       y = "Density") +
  theme_minimal()
```

## Explanation

```         
•   A density curve shows the probability distribution of the number of seeds.
```

## 3. Frequency Distribution of Fruit Diameter

```{r}
#| label: Frequency for FrDiameter
ggplot(data, aes(x = FrDiam)) +
  geom_histogram(binwidth = 2, fill = "green", color = "black") +
  labs(title = "Frequency Distribution of Fruit Diameter",
       x = "Fruit Diameter (mm)",
       y = "Frequency") +
  theme_minimal()
```

## Explanation

```         
•   A histogram shows the frequency distribution of fruit diameter in mm.

•   Binwidth is set to 2 for clarity.
```

## 4. Density Curve of Fruit Diameter

```{r}
#| label: Density FrDiameter
ggplot(data, aes(x = FrDiam)) +
  geom_density(fill = "lightgreen", alpha = 0.7) +
  labs(title = "Density Curve of Fruit Diameter",
       x = "Fruit Diameter (mm)",
       y = "Density") +
  theme_minimal()
```

## Explanation

```         
•   A density curve shows the probability distribution of fruit diameter.
```

## 5. Number of Seeds by Treatment

```{r}
#| label: Number of seeds (treatment)
ggplot(data, aes(x = Nseeds, fill = Treatment)) +
  geom_histogram(binwidth = 5, color = "black", alpha = 0.7) +
  facet_wrap(~Treatment) +
  labs(title = "Frequency Distribution of Number of Seeds by Treatment",
       x = "Number of Seeds",
       y = "Frequency") +
  theme_minimal()
```

## Explanation

```         
•   This histogram displays the number of seeds by **treatment** using **facets** to separate treatments.
```

## 6. Density Curve of Seeds by Treatment

```{r}
#| label: Density seeds (treatment)
ggplot(data, aes(x = Nseeds, color = Treatment, fill = Treatment)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~Treatment) +
  labs(title = "Density Curve of Number of Seeds by Treatment",
       x = "Number of Seeds",
       y = "Density") +
  theme_minimal()
```

## Explanation

```         
•   Density plots show the seed distribution across treatments, with transparency for overlapping curves.
```

## 7. Chi-Squared Test: Seeded vs Seedless by Treatment

```{r}
#| label: Chi-square (seed vs seedless)
# Categorize fruits as Seeded or Seedless
data <- data %>%
  mutate(Seeded = ifelse(Nseeds > 0, "Seeded", "Seedless"))

# Create a contingency table
contingency_table <- table(data$Seeded, data$Treatment)

# Calculate percentages
percentage_table <- prop.table(contingency_table, margin = 2) * 100

# Perform Chi-squared Test
chi_squared_result <- chisq.test(contingency_table)

# Print results
cat("Contingency Table:\n")
print(contingency_table)
cat("\nPercentage Table:\n")
print(round(percentage_table, 2))
cat("\nChi-squared Statistic (X^2):", chi_squared_result$statistic, "\n")
cat("P-value:", chi_squared_result$p.value, "\n")
```

## Explanation

```         
1.  Fruits are categorized as **Seeded** (Nseeds \> 0) or **Seedless**.

2.  A contingency table and percentages are generated for **Seeded vs. Seedless** across treatments.

3.  The **Chi-squared test** checks for significant associations between seed presence and treatment.
```

## 8. Percentage of Seed Presence by Treatment

```{r}
#| label: Percentage of seeds presence
#chi-squared test to compare seeded versus seedless fruit 
# Step 1: Create a new column for Seeded and Seedless categorization
data <- data %>%
  mutate(Seeded = ifelse(Nseeds > 0, "Seeded", "Seedless"))  # Categorize fruits

# Step 2: Create a contingency table for Seeded vs Seedless fruits by Treatment
contingency_table <- table(data$Seeded, data$Treatment)

# Step 3: Calculate percentages
percentage_table <- prop.table(contingency_table, margin = 2) * 100  # Percentage by Treatment

# Step 4: Run the Chi-Squared Test
chi_squared_result <- chisq.test(contingency_table)

# Step 5: Extract Chi-squared statistic and p-value
x2_value <- chi_squared_result$statistic
p_value <- chi_squared_result$p.value

# Step 6: Print results
cat("Contingency Table:\n")
print(contingency_table)
cat("\nPercentage Table (Seeded vs Seedless by Treatment):\n")
print(round(percentage_table, 2))  # Rounded for clarity
cat("\nChi-squared Statistic (X^2):", x2_value, "\n")
cat("P-value:", p_value, "\n")

#Figure 5 Plot
# Load required libraries
library(ggplot2)
library(dplyr)

# Step 1: Exclude NA values in key columns
data <- data %>%
  filter(!is.na(Nseeds), !is.na(FrDiam), !is.na(Treatment))  # Remove rows with NA values

# Step 2: Categorize Seed Presence
data <- data %>%
  mutate(SeedPresence = ifelse(Nseeds > 0, "TRUE", "FALSE"))  # Categorize as Seeded or Seedless

# Step 3: Check and force Treatment levels
data$Treatment <- factor(data$Treatment, 
                         levels = c("B", "C", "D", "E", "A"))  # Order: C-, CuGA, ZnRep, CapRep, C+

# Debugging Step: Print treatment levels
print("Treatment levels after ordering:")
print(levels(data$Treatment))

# Step 4: Calculate Percentages by Treatment and SeedPresence
percent_data <- data %>%
  group_by(Treatment, SeedPresence) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Treatment) %>%
  mutate(Percent = (Count / sum(Count)) * 100)

# Debugging Step: Print percent_data
print("Summarized percentage data:")
print(percent_data)

# Step 5: Force the graph to output
ggplot(percent_data, aes(x = Treatment, y = Percent, fill = SeedPresence)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", Percent)),
            position = position_stack(vjust = 0.5), color = "white", size = 4) +
  labs(title = "Percentage of Seed Presence by Treatment",
       x = "Treatment",
       y = "Percentage",
       fill = "Seed Presence") +
  theme_minimal() +
  scale_x_discrete(labels = c("C-", "CuGA", "ZnRep", "CapRep", "C+"))  # Correct display labels
```

## Explanation

```         
•   Percentages of **Seeded** and **Seedless** fruits are calculated for each treatment.

•   A **stacked bar chart** visualizes these percentages, with labels added for clarity.
```

## Conclusion

This analysis examined seed count, fruit diameter, and treatment effects on fruit development. The Chi-squared test evaluated the relationship between treatments and seed presence, and visualizations provided deeper insights into distributions.
