---
title: "Analysis of seed and fruit data"
author: "Arfa Ghias, Ana Belem, Cann Doan, Fatima Tahir"
format: html
editor: visual
---

```{r setup, include=FALSE}
#install.packages("dplyr")
#install.packages("knitr")
#install.packages("ggplot2")
#install.packages("agricolae")


library(dplyr)
library(knitr)
library(ggplot2)
library(agricolae)

#Folder for results
knitr::opts_chunk$set(fig.path = 'results/')
```

## Introduction

Obtaining citrus fruits without seeds is a recurrent objective for farmers as it is one of the most valued characteristics, especially in mandarins. Several agronomic techniques have been suggested to control presence of seeds, such as covering with nets and copper sulfate (CuSO4) and gibberellic acid (GA3) treatments. Natural bee repellents are also proposed to reduce the number of seeds per fruit. This study aimed to **compare the effect of several agronomic treatments to reduce the seed number** in 'Afourer' mandarin. 

The data is based on the research presented in **Scientia Horticulturae** (Volume 246, 27 February 2019, Pages 437-447), titled **“*Insect repellent and chemical agronomic treatments to reduce seed number in ‘Afourer’ mandarin: Effect on yield and fruit diameter”****.* The dataset includes measurements related to:

1.  Application of insect repellents and chemical treatments.
2.  Seed numbers in ‘Afourer’ mandarin.
3.  Yield and fruit diameter metrics.

**Objective of Article:** 

Focuses on studying methods to manage seed production in Afourer mandarins, a variety known for its desirable seedless characteristics, but it can develop seeds under certain conditions due to cross-pollination.

**Objective of Our Project:**

A. Replicate statistical analyses from the research article using available data to exercise R skills learned in class. 

B. Verify results presented in the article for seed-related data to confirm accuracy, consistency, and explore potential alternatives in processing and presenting data. 

C. To extract data independently from an open resource as opposed to a pre-defined datasets. 

## Methods

The data used in this project originates from the research article "Insect repellent and chemical agronomic treatments to reduce seed number in ‘Afourer’ mandarin: Effect on yield and fruit diameter" published in Scientia Horticulturae (2019). The original data was collected from experiments investigating the effects of different insect repellents and chemical agronomic treatments on seed production, yield, and fruit diameter in ‘Afourer’ mandarin.

Data Structure:

The data includes several variables, such as:

1.  Seed count per fruit under various treatments.
2.  Yield and fruit diameter measurements.
3.  The application of insect repellents and chemical treatments.
4.  Data from different experimental blocks and treatment groups.

The dataset is organized in Excel format, containing raw data and variables related to each of the above parameters.Our main focus was on number of seeds or seed count per fruit under various treatments.

## A. Descriptive Statistics

```{r setup, include=FALSE}
# Load necessary libraries
library(ggplot2)
knitr::opts_chunk$set(fig.path = 'results/')
# Load the data (replace 'path/to/frutasysemilla.csv' with the correct path or GitHub URL)
data <- read.csv("DATA/frutasysemillas.csv")

# Descriptive Statistics for Nseed
nseeds_mean <- mean(data$Nseeds, na.rm = TRUE)
nseeds_se <- sd(data$Nseeds, na.rm = TRUE) / sqrt(sum(!is.na(data$Nseeds)))

# Descriptive Statistics for FrDiam
frdiam_mean <- mean(data$FrDiam, na.rm = TRUE)
frdiam_se <- sd(data$FrDiam, na.rm = TRUE) / sqrt(sum(!is.na(data$FrDiam)))

# Print results
cat("Nseeds: Mean =", nseeds_mean, ", SE =", nseeds_se, "\n")
cat("FrDiam: Mean =", frdiam_mean, ", SE =", frdiam_se, "\n")
```

### Explanation:

-   **Mean**: We calculate the mean for `Nseeds` (Seed Count) and `FrDiam` (Fruit Diameter).

-   **Standard Error (SE)**: The standard error is computed by dividing the standard deviation by the square root of the sample size. This helps estimate the precision of the mean.

### Frequency Distribution and Density Curve for `Nseeds`

Next, we visualize the frequency distribution and density curve for the `Nseeds` variable to understand its distribution.

```{r}
#| label: Frequency Distribution and Density Curve
# Frequency Distribution and Density Curve for Nseeds
ggplot(data, aes(x = Nseeds)) +
  geom_histogram(aes(y = ..density..), binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +
  ggtitle("Frequency Distribution and Density Curve: Nseeds") +
  xlab("Nseeds") +
  ylab("Density") +
  theme_minimal()
```

### Explanation:

-   **Histogram**: The histogram represents the frequency distribution of `Nseeds`, with the density curve overlaid to show the estimated probability density of the data.

-   **Density Curve**: The red curve provides a smoothed representation of the distribution of `Nseeds`.

## Frequency Distribution and Density Curve for `FrDiam`

In this section, we created a similar plot for the `FrDiam` (Fruit Diameter) variable.

```{r}
#| label: Frequency FrDiam
# Frequency Distribution and Density Curve for FrDiam
ggplot(data, aes(x = FrDiam)) +
  geom_histogram(aes(y = ..density..), binwidth = 2, fill = "lightgreen", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +
  ggtitle("Frequency Distribution and Density Curve: FrDiam") +
  xlab("FrDiam") +
  ylab("Density") +
  theme_minimal()
```

### Explanation:

-   **Histogram**: This histogram represents the distribution of the `FrDiam` variable, with density plotted over it.

-   **Density Curve**: The density curve helps to visualize the smooth shape of the distribution for `FrDiam`.

## **B.** Correlation and Regression Analysis

Here, we explored the relationship between `Nseeds` and `FrDiam` using correlation and regression analysis.

### Correlation Analysis

```{r}
# Correlation Analysis between Nseeds and FrDiam
correlation <- cor(data$FrDiam, data$Nseeds, use = "complete.obs", method = "pearson")
cat("Correlation between Fruit Diameter and Seed Count:", correlation, "\n")
```

-   **Correlation**: We compute the Pearson correlation between `Nseeds` and `FrDiam`, which measures the strength and direction of the linear relationship between the two variables.

### Scatterplot with Linear Fit

```{r}
#| label: Plot FrDiam vs Nseeds
# Scatterplot with Linear Fit
ggplot(data, aes(x = FrDiam, y = Nseeds)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(title = "Scatterplot with Linear Fit: Seed Count vs Fruit Diameter",
       x = "Fruit Diameter",
       y = "Seed Count") +
  theme_minimal()
```

### Explanation:

-   **Scatterplot**: The scatterplot shows the relationship between `FrDiam` (Fruit Diameter) and `Nseeds` (Seed Count).

-   **Linear Fit**: The red line represents a linear regression model fit to the data, indicating the trend between the two variables.

### Linear Regression Model

```{r}
# Linear Regression Model
lm_model <- lm(Nseeds ~ FrDiam, data = data)
summary(lm_model)
```

### Explanation:

-   **Linear Regression**: We fit a linear regression model to predict `Nseeds` based on `FrDiam`. The model summary provides details like the coefficients, R-squared value, and p-value.

### Diagnostic Plots for Regression

```{r}
#| label: Linear model
# Plot diagnostic plots for regression model
par(mfrow = c(2, 2))  # Set plotting area
plot(lm_model)
```

### Explanation:

-   **Diagnostic Plots**: These plots are checked the assumptions of the linear regression model, such as homoscedasticity (constant variance), normality of residuals, and linearity.

## **C. Categorical** Analysis

## Chi-Squared Test: Seeded vs Seedless by Treatment

```{r}
#| label: Chi-square (seed vs seedless)
# Categorize fruits as Seeded or Seedless
data <- data %>%
  mutate(Seeded = ifelse(Nseeds > 0, "Seeded", "Seedless"))

# Create a contingency table
contingency_table <- table(data$Seeded, data$Treatment)

# Calculate percentages
percentage_table <- prop.table(contingency_table, margin = 2) * 100

# Perform Chi-squared Test
chi_squared_result <- chisq.test(contingency_table)

# Print results
cat("Contingency Table:\n")
print(contingency_table)
cat("\nPercentage Table:\n")
print(round(percentage_table, 2))
cat("\nChi-squared Statistic (X^2):", chi_squared_result$statistic, "\n")
cat("P-value:", chi_squared_result$p.value, "\n")
```

## **Explanation:**

The contingency table summarizes the counts of seeded and seedless fruits across five treatments (A, B, C, D, E):

Treatment B: Produces only seedless fruits (100% seedless, 0% seeded). This suggests that Treatment B is the most effective at preventing seed formation. Treatment C: A majority of the fruits are seedless (74.07%), though a small proportion (25.93%) still have seeds. Treatments A, D, and E: These treatments show a more balanced split between seeded and seedless fruits, with seedless fruits accounting for around 51–52% of the total.

The percentage table provides the proportion of seeded and seedless fruits within each treatment:

Treatment B stands out with 100% seedless fruits, clearly demonstrating its superior effectiveness. Treatments A, D, and E have a nearly even split between seeded and seedless fruits, indicating lower effectiveness in producing seedless fruits compared to B and C.

The chi-squared test statistic is 106.0561, indicating a large deviation between the observed and expected frequencies under the null hypothesis. A high chi-square value suggests that the distribution of seeded and seedless fruits varies significantly across treatments. The p-value is 5.04 × 10⁻²², which is extremely small (essentially zero). A p-value this small strongly rejects the null hypothesis. This indicates that the differences in the proportions of seeded and seedless fruits across treatments are statistically significant.

Treatment B is the most effective at producing seedless fruits, followed by Treatment C. Treatments A, D, and E are less effective, with more balanced proportions of seeded and seedless fruits. The chi-square test confirms that the observed differences in seed presence across treatments are not due to random chance. Treatment B is likely the best candidate for applications where seedlessness is a priority. Treatments C, D, and E might still be useful, but with less consistent results.

## 8. Percentage of Seed Presence by Treatment

```{r}
#| label: Percentage of seeds presence
#chi-squared test to compare seeded versus seedless fruit 
# Step 1: Create a new column for Seeded and Seedless categorization
data <- data %>%
  mutate(Seeded = ifelse(Nseeds > 0, "Seeded", "Seedless"))  # Categorize fruits

# Step 2: Create a contingency table for Seeded vs Seedless fruits by Treatment
contingency_table <- table(data$Seeded, data$Treatment)

# Step 3: Calculate percentages
percentage_table <- prop.table(contingency_table, margin = 2) * 100  # Percentage by Treatment

# Step 4: Run the Chi-Squared Test
chi_squared_result <- chisq.test(contingency_table)

# Step 5: Extract Chi-squared statistic and p-value
x2_value <- chi_squared_result$statistic
p_value <- chi_squared_result$p.value

# Step 6: Print results
cat("Contingency Table:\n")
print(contingency_table)
cat("\nPercentage Table (Seeded vs Seedless by Treatment):\n")
print(round(percentage_table, 2))  # Rounded for clarity
cat("\nChi-squared Statistic (X^2):", x2_value, "\n")
cat("P-value:", p_value, "\n")

#Figure 5 Plot
# Load required libraries
library(ggplot2)
library(dplyr)

# Step 1: Exclude NA values in key columns
data <- data %>%
  filter(!is.na(Nseeds), !is.na(FrDiam), !is.na(Treatment))  # Remove rows with NA values

# Step 2: Categorize Seed Presence
data <- data %>%
  mutate(SeedPresence = ifelse(Nseeds > 0, "TRUE", "FALSE"))  # Categorize as Seeded or Seedless

# Step 3: Check and force Treatment levels
data$Treatment <- factor(data$Treatment, 
                         levels = c("B", "C", "D", "E", "A"))  # Order: C-, CuGA, ZnRep, CapRep, C+

# Debugging Step: Print treatment levels
print("Treatment levels after ordering:")
print(levels(data$Treatment))

# Step 4: Calculate Percentages by Treatment and SeedPresence
percent_data <- data %>%
  group_by(Treatment, SeedPresence) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Treatment) %>%
  mutate(Percent = (Count / sum(Count)) * 100)

# Debugging Step: Print percent_data
print("Summarized percentage data:")
print(percent_data)

# Step 5: Force the graph to output
ggplot(percent_data, aes(x = Treatment, y = Percent, fill = SeedPresence)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", Percent)),
            position = position_stack(vjust = 0.5), color = "white", size = 4) +
  labs(title = "Percentage of Seed Presence by Treatment",
       x = "Treatment",
       y = "Percentage",
       fill = "Seed Presence") +
  theme_minimal() +
  scale_x_discrete(labels = c("C-", "CuGA", "ZnRep", "CapRep", "C+"))  # Correct display labels
```

## Explanation

This stacked bar chart illustrates the Percentage of Seed Presence by Treatment, comparing the proportion of seeded (TRUE) and seedless (FALSE) fruits across five treatments.

Treatment B (C-, Negative Control): 100% seedless fruits were observed, as expected for the negative control.

Treatment CuGA: 25.9% of fruits are seeded, while 74.1% are seedless. This treatment demonstrates strong suppression of seed development but is less effective than the control in achieving full seedlessness.

Treatment ZnRep: A near-even split, with 47.6% seeded and 52.4% seedless fruits. This treatment shows moderate effectiveness, with seed suppression comparable to a coin toss.

Treatment CapRep: Similarly, 48.6% seeded and 51.4% seedless. The results are almost identical to ZnRep, suggesting these two treatments have similar impacts on seed suppression.

Treatment C+: 49.0% seeded and 51.0% seedless. This treatment is also comparable to ZnRep and CapRep in its ability to produce seedless fruits.

Effectiveness of Treatments: CuGA stands out as the most effective active treatment, with the highest proportion of seedless fruits (74.1%). ZnRep, CapRep, and C+ perform similarly, with roughly equal proportions of seeded and seedless fruits, indicating lower effectiveness compared to CuGA.

## Conclusion

This analysis examined seed count, fruit diameter, and treatment effects on fruit development. The Chi-squared test evaluated the relationship between treatments and seed presence, and visualizations provided deeper insights into distributions.

## **D. Inferential statistics**

## ANOVA

ANOVAs were used to compare the mean values between treatments and blocks.

```{r}
#| label: ANOVA
# dir("DATA")
data.seeds <- read.csv("DATA/frutasysemillas.csv")

# Remove rows with missing Nseeds
data_clean <- na.omit(data.seeds[, c("Treatment", "Block", "Nseeds")])

# Perform ANOVA
anova_result <- aov(Nseeds ~ Treatment + Block, data = data_clean)
summary(anova_result)


```

Given the past results we can tell the following things regarding:

1.  **Treatment**:

    -   The effect of the treatments (CuGA, ZnRep, CapRep) on the response variable (**Nseeds**) is **significant**.

    -   The very small p-value (**2.91e-12**) tells us that **at least one treatment group is different** from the others.

2.  **Block**:

    -   The "Block" factor (replicates) also has a **significant effect** on the number of seeds. This means the blocking variable explains some of the variation in your data.

3.  **Residuals**:

    -   Residuals represent the **unexplained variability** after accounting for treatments and blocks.

#### **Conclusion**:

-   Both the **treatment** and the **block** factors influence the response variable (**Nseeds**).

-   Since the treatment effect is significant, we performed a **post-hoc test** ( Tukey's test) to find out **which treatment groups differ** from each other.

### General effect of blocks on number of seeds

```{r}
ggplot(data_clean, aes(x = Block, y = Nseeds, fill = Block)) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(title = "Effect of Blocks on Number of Seeds",
       x = "Block",
       y = "Number of Seeds") +
  scale_fill_brewer(palette = "Set3")
```

### Interaction between blocks and treatments

```{r}
interaction.plot(x.factor = data_clean$Treatment, 
                 trace.factor = data_clean$Block, 
                 response = data_clean$Nseeds, 
                 col = rainbow(length(unique(data_clean$Block))), 
                 legend = TRUE,
                 xlab = "Treatment", 
                 ylab = "Mean Number of Seeds",
                 main = "Interaction Between Treatment and Block")
```

## Tukey test

The **Tukey's post-hoc test** results helps to identify which **pairs of groups** (treatments and blocks) are significantly different after running the ANOVA.

```{r}
tukey_result <- TukeyHSD(anova_result)
tukey_result
```

### **Interpretation of Treatments**:

-   **Significant Differences** (p \< 0.05):

    -   **B-A**: Group B has a significantly lower mean than Group A.

    -   **C-B, D-B, E-B**: Groups C, D, and E have significantly higher means than Group B.

    -   **D-C and E-C**: Groups D and E have significantly higher means than Group C.

-   **Non-Significant Differences**:

    -   **C-A, D-A, E-A, and E-D**: These groups do not show significant differences.

        ## **HSD test**

The HSD test is needed to **clearly group** the treatments based on performance and identify which are significantly different.

```{r}
agricolae_result <- HSD.test(anova_result, "Treatment", group = TRUE)
print(agricolae_result$groups)
```

-   **Negative Control (B):**

    -   The mean for B is **0 (lowest group)** and is significantly different from all other treatments.

    -   This confirms that **no intervention (B)** results in **no seed production**, as expected.

-   **Positive Control (A):**

    -   The mean for A is **0.986** and belongs to group **"ab"**, indicating it performs **significantly better than the negative control (B)**.

    -   However, it is **not significantly better than treatments E and D**, meaning these two treatments perform as well as (or slightly better than) the positive control.

## Shapiro Test

The Shapiro-Wilk test is a formal statistical test for normality. The null hypothesis (**H0**) is that the data is **normally distributed**.

```{r}
#| label: Shapiro test
group_counts_nseeds <- data_clean %>%
  group_by(Treatment) %>%
  summarise(unique_count = n_distinct(Nseeds))

# Filter out treatments with only one unique value in 'Nseeds'
data_filtered_nseeds <- data_clean %>%
  filter(Treatment %in% group_counts_nseeds$Treatment[group_counts_nseeds$unique_count > 1])

# Perform the Shapiro-Wilk test on filtered data
shapiro_result_nseeds <- data_filtered_nseeds %>%
  group_by(Treatment) %>%
  summarise(p_value = shapiro.test(Nseeds)$p.value)

# Print the results
print(shapiro_result_nseeds)
# Plot the histogram and QQ plot again
ggplot(data_filtered_nseeds, aes(x = Nseeds)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  facet_wrap(~ Treatment) +
  theme_minimal() +
  labs(title = "Histogram of Number of Seeds",
       x = "Number of Seeds (Nseeds)",
       y = "Frequency")

ggplot(data_filtered_nseeds, aes(sample = Nseeds)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ Treatment) +
  theme_minimal() +
  labs(title = "QQ Plot by Treatment")
#This plots show whether this analysis follows Normality.
#The graphs show non-normal distribution
```

A **p-value \< 0.05** means we reject the null hypothesis and conclude the data is **not normally distributed**.

In this case the results let us interpret that:

-   **A**: p = 3.03e-14 → **Reject H₀** (not normal)

-   **C**: p = 1.27e-23 → **Reject H₀** (not normal)

-   **D**: p = 3.88e-15 → **Reject H₀** (not normal)

-   **E**: p = 1.75e-15 → **Reject H₀** (not normal)

Since the normality assumption is violated, we needed to use **non-parametric** alternatives like the **Kruskal-Wallis test** instead of ANOVA. The Kruskal-Wallis test does not assume normality.

## Kruskal-Wallis Test

**Due to lack of normality**, nonparametric methods were selected to compare the means among treatments and blocks by a Kruskal-Wallis rank sum test. 

```{r}
#| label: Kruskal-Wallis
tukey_result <- HSD.test(anova_result, "Treatment", group = TRUE)
tukey_groups <- data.frame(
  Treatment = rownames(tukey_result$groups),
  Group = tukey_result$groups$groups
)

# Calculate means and standard errors for each treatment
summary_data <- data.seeds %>%
  group_by(Treatment) %>%
  summarise(
    mean = mean(Nseeds, na.rm = TRUE),
    se = sd(Nseeds, na.rm = TRUE) / sqrt(n())
  )
  #The standard error measures the precision of the sample mean as an estimate of the population mean.
  #Without dividing by the sqrt the result would simply be the standard deviation, 
  #which measures the spread of individual data points, not the precision of the group mean.


# Merge Tukey groups with summary data
summary_data <- left_join(summary_data, tukey_groups, by = "Treatment")


summary_data$Treatment <- factor(summary_data$Treatment, levels = c("B", "C", "D", "E", "A"))

# Define custom colors for the bars
custom_colors <- c(
  "B" = "#1f77b4",  # Blue
  "C" = "#ff7f0e",  # Orange
  "D" = "#2ca02c",  # Green
  "E" = "#d62728",  # Red
  "A" = "#9467bd"   # Purple
)

# Define custom labels for the x-axis
x_axis_labels <- c(
  "(-) Control (B)",
  "CuGA (C)",
  "ZnRep (D)",
  "CapRep (E)",
  "+ Control (A)"
)
# Create the bar plot with error bars and Tukey groups
ggplot(summary_data, aes(x = Treatment, y = mean, fill = Treatment)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_text(aes(label = Group, y = mean + se + 0.1), vjust = 0) +
  labs(title = "Mean Number of Seeds by Treatment",
       x = "Treatment",
       y = "Mean Number of Seeds",
       fill = "Treatments") +
  theme_minimal() +
  scale_fill_manual(values = custom_colors) +  # Apply custom bar colors
  scale_x_discrete(labels = x_axis_labels) +  # Use custom x-axis labels
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),  # Rotate x-axis labels
        legend.position = "right")

```

According to the given graph:

-   **The best option is the negative control (B)**, as it produced the **fewest seeds**.

-   **CuGA (C)** could also be useful, as it significantly reduced seed count, although it’s not as effective as the **negative control**.

-   **ZnRep (D)** and **CapRep (E)** are the least desirable treatments since they resulted in the **highest seed production**.

Thus, we should prioritize treatments that either have **no seeds (B)** or minimize seed count, such as **CuGA (C)**, to achieve fruits with the fewest seeds.

## E. Visualization Techniques

In this section, we visualize the seed counts by treatment using various types of plots.

### Box Plot by Treatment

```{r}
#| label: Boxplot Treatments
# Load necessary libraries
library(ggplot2)

# Ensure Treatment is a factor and ordered correctly
data$Treatment <- factor(data$Treatment, levels = c("A", "B", "C", "D", "E"))

# Define colors for each treatment
treatment_colors <- c("A" = "#1f77b4",  # Blue for Negative Control
                      "B" = "#ff7f0e",  # Orange for Variable 1
                      "C" = "#2ca02c",  # Green for Variable 2
                      "D" = "#d62728",  # Red for Variable 3
                      "E" = "#9467bd")  # Purple for Positive Control

# Create the box plot
ggplot(data, aes(x = Treatment, y = Nseeds, fill = Treatment)) +
  geom_boxplot() +  # Draw box plots
  scale_fill_manual(values = treatment_colors) +  # Assign colors
  labs(title = "Box Plot of Seed Counts by Treatment",
       x = "Treatment",
       y = "Seed Count") +
  theme_minimal()
```

### Explanation:

-   **Box Plot**: The box plot visualizes the distribution of `Nseeds` for each treatment, showing the median, quartiles, and outliers. Each treatment is colored differently for clarity.

    ## **Bar Chart with Error Bars**

    ```{r}
    # Load necessary libraries
    library(dplyr)

    # Calculate means and confidence intervals
    summary_data <- data %>%
      group_by(Treatment) %>%
      summarise(Mean = mean(Nseeds, na.rm = TRUE),
                SE = sd(Nseeds, na.rm = TRUE) / sqrt(n()),
                Lower = Mean - qt(1 - 0.05 / 2, n() - 1) * SE,
                Upper = Mean + qt(1 - 0.05 / 2, n() - 1) * SE)

    # View the summary data
    print(summary_data)
    ```

### Explanation:

-   **Summary Statistics**: We calculate the mean and standard error (SE) for `Nseeds` by treatment, along with the confidence intervals (Lower and Upper).

-   **Error Bars**: These represent the uncertainty around the mean seed count for each treatment.

```{r}
#| label: Standard Error Bars
# Bar Chart with Error Bars
ggplot(summary_data, aes(x = Treatment, y = Mean, fill = Treatment)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2) +
  labs(title = "Mean Seed Count per Treatment with Confidence Intervals",
       x = "Treatment",
       y = "Mean Seed Count") +
  theme_minimal()
```

### Explanation:

-   **Bar Chart**: The bar chart displays the mean seed count for each treatment with error bars representing the 95% confidence interval.

### Scatterplot of Seed Count vs Fruit Diameter

```{r}
#| label: Seed Count vs Fruit Diameter
# Scatterplot
ggplot(data, aes(x = FrDiam, y = Nseeds)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Scatterplot of Seed Count vs Fruit Diameter",
       x = "Fruit Diameter",
       y = "Seed Count") +
  theme_minimal()
```

### Explanation:

-   **Scatterplot**: This scatterplot shows the relationship between `FrDiam` and `Nseeds`, with a red linear regression line.

### Violin Plot of Seed Counts by Treatment

```{r}
{r}
#| label: Seed Count Violin
# Violin Plot
ggplot(data, aes(x = Treatment, y = Nseeds, fill = Treatment)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white", outlier.color = "red") +
  labs(title = "Violin Plot of Seed Counts by Treatment",
       x = "Treatment",
       y = "Seed Count") +
  theme_minimal()
```

### Explanation:

-   **Violin Plot**: The violin plot visualizes the distribution of `Nseeds` for each treatment, showing the kernel density estimate along with a box plot inside the violin to represent summary statistics.

## **Recommendation for Visualizations:**

-   For understanding the overall distribution of data, **density plots** are effective as they highlight the shape of the distribution better.
-   For comparing treatments, **box plots** and **violin plots** clearly show variations and central tendencies.
-   **Bar charts with error bars** are ideal for summarizing mean values with their associated uncertainties.
