---
title: "Less Than Two Seeds"
author: "Ana Belem García, Arfa Ghias, Cann Doan, Fatima Tahir"
format: html
editor: visual
---

```{r setup, include=FALSE}
#install.packages("dplyr")
#install.packages("knitr")
#install.packages("ggplot2")
#install.packages("agricolae")


library(dplyr)
library(knitr)
library(ggplot2)
library(agricolae)

#Folder for results
knitr::opts_chunk$set(fig.path = 'results/')
```

## Analysis with less than two seeds fruits

In the fruit industry, seedless fruits or fruits with very few seeds are highly sought after for their consumer appeal, as they offer greater convenience and are often preferred in various culinary applications. While the ideal is to produce entirely seedless fruits, those with a minimal seed count are still considered desirable. Therefore, we decided to reanalyze our data with a new focus: treating fruits with **2 or fewer seeds** as **seedless**. This approach allows for a more refined understanding of the treatments that effectively minimize seed production, which is crucial for enhancing the marketability of fruit products.

```{r}
# Load data
#| label: TwoSeedsPercen
data <- read.csv("DATA/frutasysemillas.csv")
data <- na.omit(data)
data <- data %>%
  mutate(SeedPresence = ifelse(Nseeds > 2, "TRUE", "FALSE"))  # Categorize as Seeded or Seedless

# Step 3: Check and force Treatment levels
data$Treatment <- factor(data$Treatment, 
                         levels = c("B", "C", "D", "E", "A"))  # Order: C-, CuGA, ZnRep, CapRep, C+

# Debugging Step: Print treatment levels
#print("Treatment levels after ordering:")
#print(levels(data$Treatment))

# Step 4: Calculate Percentages by Treatment and SeedPresence
percent_data <- data %>%
  group_by(Treatment, SeedPresence) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Treatment) %>%
  mutate(Percent = (Count / sum(Count)) * 100)

# Debugging Step: Print percent_data
#print("Summarized percentage data:")
#print(percent_data)

# Step 5: Force the graph to output
ggplot(percent_data, aes(x = Treatment, y = Percent, fill = SeedPresence)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", Percent)),
            position = position_stack(vjust = 0.5), color = "white", size = 4) +
  labs(title = "Percentage of Seed Presence by Treatment",
       x = "Treatment",
       y = "Percentage",
       fill = "Seed Presence") +
  theme_minimal() +
  scale_x_discrete(labels = c("C-", "CuGA", "ZnRep", "CapRep", "C+"))  # Correct display labels


```

```{r}
# Read and clean the data
data <- read.csv("DATA/frutasysemillas.csv", stringsAsFactors = TRUE)
data_clean <- na.omit(data[, c("Treatment", "Tree", "Nseeds")])

# Classify fruits as seedless or seeded
data_clean <- data_clean %>%
  mutate(SeedStatus = ifelse(Nseeds <= 2, "Seedless", "Seeded"))

# Create a contingency table for Treatment and SeedStatus
contingency_table <- table(data_clean$Treatment, data_clean$SeedStatus)
print(contingency_table)

# Perform Fisher's Exact Test
fisher_result <- fisher.test(contingency_table, workspace = 2e7) # Increase workspace
print(fisher_result)

# Reorder treatments for plotting
data_clean$Treatment <- factor(data_clean$Treatment, levels = c("B", "C", "D", "E", "A"))

# Calculate proportions of Seedless and Seeded fruits for each Treatment
proportion_data <- data_clean %>%
  group_by(Treatment, SeedStatus) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Proportion = Count / sum(Count))

# Perform pairwise Fisher's tests manually
pairwise_results <- combn(levels(data_clean$Treatment), 2, function(pair) {
  sub_table <- contingency_table[pair, ]
  fisher_res <- fisher.test(sub_table)
  data.frame(
    Group1 = pair[1],
    Group2 = pair[2],
    p_value = fisher_res$p.value
  )
}, simplify = FALSE) %>%
  bind_rows()

# Adjust p-values for multiple comparisons
pairwise_results <- pairwise_results %>%
  mutate(adjusted_p_value = p.adjust(p_value, method = "bonferroni"),
         significance = case_when(
           adjusted_p_value > 0.05 ~ "ns",
           adjusted_p_value <= 0.001 ~ "***",
           adjusted_p_value <= 0.01 ~ "**",
           adjusted_p_value <= 0.05 ~ "*"
         ))

print(pairwise_results)

# Add annotations for the plot using both the p-value and corresponding symbol
annotations <- data.frame(
  Treatment = levels(data_clean$Treatment),
  Seedless_Proportion = proportion_data %>%
    filter(SeedStatus == "Seedless") %>%
    pull(Proportion),
  Label = sprintf("p = %.3f %s", pairwise_results$p_value, pairwise_results$significance)  # Combining p-value and significance symbol
)

# Create a clustered bar plot with both p-values and significance symbols
# Create a clustered bar plot with both p-values and significance symbols
ggplot(proportion_data, aes(x = Treatment, y = Proportion, fill = SeedStatus)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), alpha = 0.8) +
  # Add text labels for p-values and significance symbols only for Seedless category
  geom_text(data = annotations, aes(x = Treatment, y = Seedless_Proportion, label = Label), 
            inherit.aes = FALSE, position = position_dodge(width = 0.8), 
            vjust = -0.5, size = 3.5, color = "black") +  # Adjust vjust to move text upwards
  scale_fill_brewer(palette = "Set2", name = "Seed Status") +
  theme_minimal() +
  labs(
    title = "Proportion of Seedless and Seeded Fruits by Treatment",
    x = "Treatment",
    y = "Proportion"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(size = 10),  # Adjust axis text size for better readability
    axis.title = element_text(size = 12)  # Adjust axis title size for better readability
  )

```
